Porting the Steel Bank Common Lisp Compiler and Runtime to the Nintendo Switch NX Platform

### 2025-01-17 els2025-sbcl-nx/paper.tex
Some more paper work

@@

\section{Garbage Collection}\label{gc}

@@

\section{Further Work}\label{further-work}

>>>>

Patton\cite{patton2023parallel} and Schwartz\cite{schwartz2018dynamic} describe some details about SBCL's garbage collector which we modify for our port.

@@

\colorbox{red}{
  \textbf{TODO CHARLES}: Add some details here.
}

\section{Garbage Collection}\label{gc}
SBCL' default garbage collector, ``gencgc'', is a stop-the-world collector. In order to start collecting, it needs to ``park'' all threads in a safe location where they won't access any memory that the collector might move or change during collection.

On Unix systems signals are used to park the threads. By sending a signal to every other thread, each thread will enter a signal handler, which can then park the thread. This is advantageous, since it leverages an operating system mechanism to interrupt thread execution and frees the thread from checking whether it should park.

On Windows no equivalent mechanism is available, so a technique called ``safepoints'' is employed instead. The compiler injects code at strategic locations, such as at function call boundaries and loop returns, which checks whether the thread should park explicitly. This strategy is notably slower than using signals, since the safepoints are not free, and is finicky, as the compiler could inject too many safepoints (slowing execution of user code significantly) or too few safepoints (making threads take too long to park).

In order to make the safepoint test as small and cheap as possible SBCL does employ another trick on Windows: each safepoint simply writes to a memory region. When GC is triggered, this region is marked as protected, causing any thread that tries to write to it to trap, which SBCL can intercept to park the thread instead.

On the NX we cannot install trap handlers, so the current safepoint strategy fails.

\colorbox{red}{
  \textbf{TODO CHARLES}: Add some details here.
}

@@

\section{Further Work}\label{further-work}
Currently we rely on the fasteval system to circumvent runtime compilation restrictions. This is especially vital for CLOS, since the discriminating function is only compiled on first call of a generic function. However, since we know that we don't dynamically add or remove methods, we should be able to pre-compile all of these functions as well. We'd like to add such a CLOS freeze step to the build pipeline.

Callbacks from C to Lisp are currently left unimplemented. Callbacks in SBCL are implemented in part as hand-written assembly stubs, and are very hard to maintain. The current stubs for ARM64 do not work correctly in the presence of the shrinkwrapping steps that we make use of.

Finally there are a number of improvements we've made that we would like to upstream to lessen the maintenance burden. We've already contributed a bunch of the changes back upstream, but a lot of it is also tied to the proprietary SDK and cannot be open-sourced due to the NDA, and some other changes are too contentious for the rest of the SBCL team to want to upstream.
